\documentclass[12pt]{article}
\usepackage[svgnames,x11names,table]{xcolor}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{float}
\usepackage{amsmath}
\usepackage{esint}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage[thicklines]{cancel}

\hypersetup{
    colorlinks,
    citecolor=blue,
    filecolor=black,
    linkcolor=black,
    urlcolor=RoyalBlue4,
}

\title{PEU 356 Assignment 7}
\author{Mohamed Hussien El-Deeb (201900052)}
\date{\today}

\DeclareMathOperator{\sech}{sech}
\DeclareMathOperator{\csch}{csch}

\begin{document}

\maketitle
\tableofcontents
\hypersetup{linkcolor=RoyalBlue4}

\newpage
\section{Question 1}

\subsection{Problem}

Find the eigenvalues and corresponding normalized eigenvectors of the matrix

\[
    A = \begin{pmatrix}
        0 & 1 & 1 \\
        1 & 0 & 1 \\
        1 & 1 & 0 \\
    \end{pmatrix}
\]

Orthogonalize any degenerate eigenvectors.

ANS. \(\lambda = -1, -1, 2. \)

\subsection{Solution}

\newpage
\section{Question 2}

\subsection{Problem}

Find the eigenvalues and corresponding normalized eigenvectors of the matrix

\[
    A = \begin{pmatrix}
        5 & 0 & 2 \\
        0 & 1 & 0 \\
        2 & 0 & 2 \\
    \end{pmatrix}
\]

Orthogonalize any degenerate eigenvectors.

ANS. \(\lambda = 0, 0, 3. \)

\subsection{Solution}

\newpage
\section{Question 3}

\subsection{Problem}

Describe the geometric properties of the surface

\[
    x^2+2 x y+2 y^2+2 y z+z^2=1 .
\]

How is it oriented in 3-D space? Is it a conic section? If so, which kind?

\subsection{Solution}

\newpage
\section{Question 4}

\subsection{Problem}

A system of \(N\) degrees of freedoms is subject to the following potential

\[
    V=\frac{1}{2} V_{i j} x_i x_j, \quad i, j=1, \ldots, N .
\]

(i) Derive Newton's second law for the system.

(ii) Assume we are looking for solutions of the form

\[
    X(t)=X(0) \sin \omega t, \quad X(t)=\left(\begin{array}{c}
            x_1(t) \\
            \vdots \\
            x_N(t)
        \end{array}\right) .
\]

Show that \(X(0)\) satisfies an eigenvalue problem.

(iii) What are the conditions on the eigenvalues of \(V_{i j}\) in order to have normal modes (i.e., oscillatory SII motion with a fixed frequency).

\subsection{Solution}

\newpage
\section{Question 5}

\subsection{Problem}

Consider the previous example with the following potential

\[
    V=x_1 x_2+x_1 x_3+x_2 x_3 .
\]

Find the normal modes.

\subsection{Solution}



\newpage
\bibliographystyle{plain}
\bibliography{references}
\nocite{El-Deeb_PEU-356_Assignments}

\end{document}